---
title: "Non-Parametric Econometrics PS-6 Exercise 1"
author: "Ferhat Kutluca"
date: "2024-02-13"
output:
html_document: default
pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Load Libraries

```{r}

library(np)
library(MASS)

```


### Set Parameters

```{r}
# Set seed for reproducibility
set.seed(123)

# Define the sample size and number of random samples
n <- 500
S <- 100

delta <- 1
beta_hat <- numeric(S)
delta_hat <- numeric(S)

```


### Generate Data and Estimate Parameters

```{r}
for (s in 1:S) {
  Z <- runif(n)
  T <- rbinom(n, 1, prob = sin(2*Z))
  g <- 5*sin(3*Z) + Z^4
  u <- runif(n, min = -1, max = 1)
  Y <- delta*T + g + u
  
  model <- lm(Y ~ T)
  beta_hat[s] <- coef(model)["T"]
  
  E_T_Z <- npreg(txdat=Z, tydat=T, multistart = FALSE)$mean
  E_Y_Z <- npreg(txdat=Z, tydat=Y, multistart = FALSE)$mean
  model_rt <- lm(I(Y - E_Y_Z) ~ I(T - E_T_Z))
  delta_hat[s] <- coef(model_rt)["I(T - E_T_Z)"]
}


```

### Report Results

```{r}

cat("Average slope coefficient (delta estimate):", mean(beta_hat), "\n")
cat("Average delta estimate using Robinson's Transformation:", mean(delta_hat), "\n")

```
(i) The average slope coefficient (delta estimate) obtained by regressing $Y_i$ on $T_i$ ignoring the component $g(Z_i)$ is **2.378575 (S = 100)**; **2.381133 (S = 1000)**. However, this is different from the true $δ$ value of 1. This is not a good estimator of $δ(1)$. This could be a result of ignoring the $g(Z_i)$ component in the model. This component could influence the relationship between $Y_i$ and $T_i$, and therefore, ignoring this component could lead to a misleading estimate of $δ$.

(ii) The average delta estimate obtained using Robinson's Transformation is **0.9860977 (S = 100)**; **0.9900204 (S = 1000)**. This value is quite close to the true delta value of 1 and the estimator is centered correctly around $δ(1)$. This suggests that Robinson's Transformation, by accounting for the $g(Z_i)$ component in the model, provides a more accurate estimate of $δ$.

Therefore, the estimate obtained using Robinson's Transformation is more accurate than the estimate obtained by ignoring the $g(Z_i)$ component. This demonstrates the effectiveness of Robinson's Transformation in estimating partially linear models. This demonstrates the effectiveness of Robinson's Transformation in estimating partially linear models.
